---
title: Diversity indicators using OBIS data (EEZ)
date: "`r Sys.Date()`"
author: Pieter Provoost
output: (function(...) {
  rmdformats::robobook(toc_depth = 4, pandoc_args = c("+RTS", "-K2000m", "-RTS"), ...) })
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })  
---

This notebook builds on the main notebook to generate indicator maps for EEZ only.

# Read the occurrence data

```{r message=FALSE, warning=FALSE}
occ <- readRDS("occurrence.rds")
```

# Gridded maps with masked EEZs

## Create a discrete global grid

Create an ISEA discrete global grid of resolution 9 using the [dggridR](https://github.com/r-barnes/dggridR) package and assign cell numbers to the occurrence data:

```{r warning=FALSE, message=FALSE}
library(dggridR)

dggs <- dgconstruct(projection = "ISEA", topology = "HEXAGON", res = 9)
occ$cell <- dgtransform(dggs, occ$decimallatitude, occ$decimallongitude)
```

## Calculate metrics

This calculates stations per 1000 square km. Make sure to adjust if the resolution is changed.

```{r warning=FALSE, message=FALSE}
library(dplyr)

metrics <- occ %>%
  group_by(cell, decimallongitude, decimallatitude) %>%
  summarize(records = sum(records)) %>%
  group_by(cell) %>%
  summarize(records = sum(records), stations = n()) %>%
  mutate(stations = stations / 2591.4 * 1000)

metrics
```

Add cell geometries to the metrics table:

```{r warning=FALSE, message=FALSE}
library(sf)

grid <- dgearthgrid(dggs, frame = FALSE, wrapcells = FALSE)
grid$cell <- names(grid)
grid_sf <- grid %>%
  st_as_sf() %>%
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=230"))

grid_cropped <- grid_sf %>% st_crop(c(xmin = -180, ymin = -85, xmax = 180, ymax = 85))

metrics <- merge(metrics, grid_sf, by.x = "cell", by.y = "cell", all.y = TRUE) %>%
    st_sf()
```

## Load high seas shapefile

Use the high seas shapefile from https://www.marineregions.org/ to mask the high seas:

```{r warning=FALSE, message=FALSE}
highseas <- read_sf("World_High_Seas_v1_20200826/High_Seas_v1_densified.shp")
```

## Create maps

```{r warning=FALSE, message=FALSE}
library(rnaturalearth)
library(rnaturalearthdata)
library(viridis)
library(ggplot2)

world <- ne_countries(scale = "medium", returnclass = "sf")

ggplot() +
  geom_sf(data = metrics, aes_string(fill = "stations", color = "stations"), lwd = 0.1) +
  scale_fill_viridis(option = "inferno", na.value = "#000000", name = expression("OBIS: number of stations per 1000"~km^2), trans = "log10") +
  scale_color_viridis(option = "inferno", na.value = "#000000", name = expression("OBIS: number of stations per 1000"~km^2), trans = "log10") +
  geom_sf(data = world, fill = "#dddddd", color = NA) +
  geom_sf(data = highseas, fill = "#ffffff", color = NA, alpha = 0.5) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(), 
    panel.background = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "top"
  ) +
  guides(fill = guide_colourbar(barwidth = 27, barheight = NULL, title.position = "bottom", title.hjust = 0.5)) +
  xlab("") +
  ylab("") +
  coord_sf(crs = "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
```

```{r warning=FALSE, message=FALSE}
ggsave("map/map_eez.png", width = 11, height = 6, dpi = 400, scale = 2)
```

```{r warning=FALSE, message=FALSE}
ggplot() +
  geom_sf(data = metrics, aes_string(fill = "stations", color = "stations"), lwd = 0.1) +
  scale_fill_distiller(palette = "Spectral", na.value = "#3288bd", name = expression("OBIS: number of stations per 1000"~km^2), trans = "log10") +
  scale_color_distiller(palette = "Spectral", na.value = "#3288bd", name = expression("OBIS: number of stations per 1000"~km^2), trans = "log10") +
  geom_sf(data = world, fill = "#dddddd", color = NA) +
  geom_sf(data = highseas, fill = "#ffffff", color = NA, alpha = 0.5) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(), 
    panel.background = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "top"
  ) +
  guides(fill = guide_colourbar(barwidth = 27, barheight = NULL, title.position = "bottom", title.hjust = 0.5)) +
  xlab("") +
  ylab("") +
  coord_sf(crs = "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
```

```{r warning=FALSE, message=FALSE}
ggsave("map/map_eez_spectral.png", width = 11, height = 6, dpi = 400, scale = 2)
```

```{r warning=FALSE, message=FALSE}
ggplot() +
  geom_sf(data = metrics, aes_string(fill = "stations", color = "stations"), lwd = 0.1) +
  scale_fill_distiller(palette = "Spectral", na.value = "#3288bd", name = expression("OBIS: number of stations per 1000"~km^2), trans = "log10") +
  scale_color_distiller(palette = "Spectral", na.value = "#3288bd", name = expression("OBIS: number of stations per 1000"~km^2), trans = "log10") +
  geom_sf(data = world, fill = "#dddddd", color = "#000000", lwd = 0.2) +
  geom_sf(data = highseas, fill = "#ffffff", color = "#000000", alpha = 0.5, lwd = 0.2) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(), 
    panel.background = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "top"
  ) +
  guides(fill = guide_colourbar(barwidth = 27, barheight = NULL, title.position = "bottom", title.hjust = 0.5)) +
  xlab("") +
  ylab("") +
  coord_sf(crs = "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
```

```{r warning=FALSE, message=FALSE}
ggsave("map/map_eez_spectral_lines.png", width = 11, height = 6, dpi = 400, scale = 2)
```

```{r warning=FALSE, message=FALSE}
ggplot() +
  geom_sf(data = metrics, aes_string(fill = "stations", color = "stations"), lwd = 0.1) +
  scale_fill_distiller(palette = "Spectral", na.value = "#ffffff", name = expression("OBIS: number of stations per 1000"~km^2), trans = "log10") +
  scale_color_distiller(palette = "Spectral", na.value = "#ffffff", name = expression("OBIS: number of stations per 1000"~km^2), trans = "log10") +
  geom_sf(data = world, fill = "#dddddd", color = NA) +
  geom_sf(data = highseas, fill = "#ffffff", color = NA, alpha = 0.5) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(), 
    panel.background = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "top"
  ) +
  guides(fill = guide_colourbar(barwidth = 27, barheight = NULL, title.position = "bottom", title.hjust = 0.5)) +
  xlab("") +
  ylab("") +
  coord_sf(crs = "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
```

```{r warning=FALSE, message=FALSE}
ggsave("map/map_eez_spectral_white.png", width = 11, height = 6, dpi = 400, scale = 2)
```

# Metrics by EEZ

First create a table of all sampling locations:

```{r warning=FALSE, message=FALSE}
stations <- occ %>%
  group_by(decimallongitude, decimallatitude) %>%
  summarize(records = sum(records)) %>%
  st_as_sf(coords = c("decimallongitude", "decimallatitude"), crs = 4326)
write.csv(stations %>% select(decimallongitude, decimallatitude), file = "stations.csv", row.names = FALSE)
```

My intention was to count the stations per polygon in R, but for performance reasons I'm using QGIS's `Count points in polygon` instead.

```{r warning=FALSE, message=FALSE}
# work in progress

#library(furrr)
#eez <- read_sf("World_EEZ_v11_20191118/eez_v11.shp")
#future::plan(multiprocess)
#options(future.globals.maxSize=1e10)
#pip <- future_map(1:nrow(eez), function(i) {
#  ns <- sum(st_within(stations$geometry, eez$geometry[i], sparse = FALSE))
#  return(ns)
#}, .progress = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(rnaturalearth)
library(rnaturalearthdata)
library(ggplot2)
library(dplyr)
library(sf)

world <- ne_countries(scale = "medium", returnclass = "sf")
eez <- read_sf("shapefile/eez_v11_counts_simplified.gpkg")

eez <- eez %>%
  mutate(density = NUMPOINTS / AREA_KM2 * 1000) %>%
  mutate(bin = cut(density, breaks = c(0, 10, 100, 500, 1000, 11000))) %>%
  filter(!is.na(bin)) %>%
  mutate(bin = forcats::fct_recode(bin, "0 - 10" = "(0,10]", "10 - 100" = "(10,100]", "100 - 500" = "(100,500]", "500 - 1000" = "(500,1e+03]", "1000 - 11000" = "(1e+03,1.1e+04]"))

ggplot() +
  geom_sf(data = eez, aes_string(fill = "bin", color = "bin"), lwd = 0.3, color = "#000000") +
  scale_fill_brewer(palette = "Spectral", na.value = "#ffffff", name = expression("Number of stations per 1000"~km^2), direction = -1) +
  scale_color_brewer(palette = "Spectral", na.value = "#ffffff", name = expression("Number of stations per 1000"~km^2), direction = -1) +
  geom_sf(data = world, fill = "#6c6c61", color = "#000000", lwd = 0.3) +
  guides(color = guide_legend(ncol = 1), fill = guide_legend(ncol = 1)) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(), 
    panel.background = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "bottom"
  ) +
  xlab("") +
  ylab("") +
  coord_sf()
```

```{r warning=FALSE, message=FALSE}
ggsave("map/map_eez_bins.png", width = 11, height = 6, dpi = 400, scale = 2)
```
