---
title: Diversity indicators using OBIS data
date: "`r Sys.Date()`"
author: Pieter Provoost
output: (function(...) {
  rmdformats::robobook(toc_depth = 4, pandoc_args = c("+RTS", "-K2000m", "-RTS"), ...) })
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })  
---

This notebook shows how to calculate diversity indicators layers based on OBIS data. The indicators are calculated for a discrete global grid.

## Read the occurrence data

Exports of all OBIS presence records are available from https://obis.org/manual/access/. Extract the archive, read the location and species identifier columns, and summarize by location:

```{r message=FALSE, warning=FALSE}
library(readr)
library(dplyr)

if (!file.exists("occurrence.rds")) {
  occ <- read_csv("../occurrence.csv", col_type = cols_only(decimallongitude = col_double(), decimallatitude = col_double(), species = col_character())) %>%
    group_by(decimallongitude, decimallatitude, species) %>%
    summarize(records = n()) %>%
    as_tibble()
  saveRDS(occ, "occurrence.rds")
} else {
  occ <- readRDS("occurrence.rds")
}
```

## Create a discrete global grid

Create an ISEA discrete global grid of resolution 9 using the [dggridR](https://github.com/r-barnes/dggridR) package:

```{r warning=FALSE, message=FALSE}
library(dggridR)

dggs <- dgconstruct(projection = "ISEA", topology = "HEXAGON", res = 9)
```

Here's on overview of all possible resolutions:

```{r warning=FALSE}
inf <- dginfo(dggs)
```

Then assign cell numbers to the occurrence data:

```{r warning=FALSE, message=FALSE}
occ$cell <- dgtransform(dggs, occ$decimallatitude, occ$decimallongitude)
```

## Calculate metrics

The following function calculates the number of records, species richness, Simpson index, Shannon index, Hurlbert index (n = 50), and Hill numbers for each cell.

```{r warning=FALSE, message=FALSE}
library(gsl)

calc <- function(df, esn = 50) {
  t1 <- df %>%
    group_by(cell, species) %>%
    summarize(ni = sum(records))
  t2 <- t1 %>%
    group_by(cell) %>%
    mutate(n = sum(ni))
  t3 <- t2 %>%
    group_by(cell, species) %>%
    mutate(
      hi = -(ni/n*log(ni/n)),
      si = (ni/n)^2,
      qi = ni/n,
      esi = case_when(
        n-ni >= esn ~ 1-exp(lngamma(n-ni+1)+lngamma(n-esn+1)-lngamma(n-ni-esn+1)-lngamma(n+1)),
        n >= esn ~ 1
      )
    )
  t4 <- t3 %>%
    group_by(cell) %>%
    summarize(
      n = sum(ni),
      sp = n(),
      shannon = sum(hi),
      simpson = sum(si),
      maxp = max(qi),
      es = sum(esi)
    )
  result <- t4 %>%
    mutate(
      hill_1 = exp(shannon),
      hill_2 = 1/simpson,
      hill_inf = 1/maxp
    )
  return(result)
}
```

Perform the calculation on species level data:

```{r warning=FALSE, message=FALSE}
metrics <- occ %>%
  filter(!is.na(species)) %>%
  calc(50)

metrics
```

Add cell geometries to the metrics table:

```{r warning=FALSE, message=FALSE}
library(sf)

grid <- dgearthgrid(dggs, frame = FALSE, wrapcells = FALSE)
grid_sf <- grid %>%
  st_as_sf() %>%
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=230"))
grid_sf$cell <- names(grid)

metrics <- merge(metrics, grid_sf, by.x = "cell", by.y = "cell") %>%
    filter(st_intersects(geometry, st_as_sfc("SRID=4326;POLYGON((-180 85,180 85,180 -85,-180 -85,-180 85))"), sparse = FALSE))
```

Let's check the results by creating a Shannon index map:

```{r warning=FALSE, message=FALSE}
library(rnaturalearth)
library(rnaturalearthdata)
library(viridis)
library(ggplot2)

world <- ne_countries(scale = "medium", returnclass = "sf")

ggplot() +
    geom_sf(data = metrics, aes_string(fill = "shannon", geometry = "geometry"), lwd = 0) +
    scale_fill_viridis(option = "inferno", na.value = "white", name = "Shannon index") +
    geom_sf(data = world, fill = "#dddddd", color = NA) +
    theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(), panel.grid.minor.x = element_blank(), panel.grid.minor.y = element_blank(), panel.background = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank()) + xlab("") + ylab("") +
    coord_sf()
```

## Output
### Shapefile

Write the `metrics` table to a shapefile:

```{r warning=FALSE, message=FALSE, results='hide'}
if (file.exists("shapefile/indicators.shp")) file.remove("shapefile/indicators.shp")
st_write(metrics, "shapefile/indicators.shp", layer = "indicators")
files <- dir("shapefile", full.names = TRUE)
zip(zipfile = "shapefile/metrics.zip", files = files)
file.remove(files)
```

### Map image

This create a high resolution PNG map image.

```{r warning=FALSE, message=FALSE}
metrics <- st_sf(metrics, sf_column_name = "geometry")
st_crs(metrics) <- 4326

ggplot() +
  geom_sf(data = metrics, aes_string(fill = "n", color = "n", geometry = "geometry"), lwd = 0.04) +
  scale_color_viridis(option = "inferno", na.value = "white", name = "Number of records", trans = "log10") +
  scale_fill_viridis(option = "inferno", na.value = "white", name = "Number of records", trans = "log10") +
  geom_sf(data = world, fill = "#dddddd", color = "#666666", lwd = 0.1) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(), 
    panel.background = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none"
  ) +
  xlab("") + ylab("") +
  coord_sf(crs = "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs" )

ggsave("map/map.png", width = 12, height = 6, dpi = 600)
```
